### 2020-01-14 draft.tex
First fully workable commit

\title{{\Large \bf Industry Compliance Costs Under the Renewable Fuel Standard: Evidence from Compliance Credits\thanks{Early drafts of this paper benefited from comments by William Shughart, Ben Blau, and Megan Hansen. The Center for Growth and Opportunity at Utah State University provided financial support for data procurement and research. All errors are the author's own.}}}
\author{Arthur R. Wardle\footnote{
PhD Student, Department of Agricultural \& Resource Economics, University of California, Berkeley. \href{mailto:arw@berkeley.edu}{\tt arw@berkeley.ed} (corresponding author).}\\
Sherzod B. Akhundjanov\footnote{Assistant Professor, Department of Applied Economics, Utah State University. \href{matilto:sherzod.akhundjanov@usu.edu}{\tt sherzod.akhundjanov@usu.edu}}}

@@

Placeholder abstract
\newline

@@

The Renewable Fuel Standard (RFS), created under the Energy Policy Act of 2005 and greatly expanded by the Energy Independence and Security Act of 2007, mandates the use of various biofuels in domestic transportation fuels.\footnote{A complete description of how the RFS works is available in \cite{Schnepf2013}.} The statute itself includes volumetric mandates for cellulosic, biomass-based biodiesel, ``advanced," and total renewable fuels through 2022. Obligated parties (fuel refiners and importers) are required to submit specified numbers of Renewable Identification Numbers (RINs) to comply with the standard. RINs are created by biorefineries, which link them to barrels of renewable fuels, and are split from those barrels upon fuel blending. RINs can be traded, allowing obligated parties to comply with RFS mandates either by blending renewable fuels themselves or buying excess RINs from other parties. The construction of the RIN market is functionally equivalent to other market-based environmental regulations such as pollution permits.

Because regulated firms can comply with the RFS either by blending additional biofuels or by buying RINs, the basic fundamental value of a RIN (with some complications, described later in this paper) is the marginal cost to the refining sector of blending an additional unit of biofuel.

Even high RIN prices do not necessarily impact the bottom line of obligated parties if these costs are easy to pass through to consumers and demand for transportation fuels is inelastic. Indeed, a large literature establishes that refiners are able to fully (or even more than fully) pass RFS compliance costs onto consumers \citep{Burkhardt2016,Knittel2017,Pouliot2017,Li2019,Lade2019}. This, however, does not imply that the RFS has no financial impact US oil refiners whatsoever. Complete pass-through does not replace the profit margins on refined crude oil that counterfactually would  have been sold without the mandate and may impose infrastructure costs to accommodate ethanol blends, negotiating costs with biorefineries, and myriad costs of doing business not fully captured in models of cost pass-through. \cite{KnittelSmith2015} give a fuller description of ethanol's impact on oil refining profitability. The goal of our analysis is to establish how changes in the prices of RFS compliance credits (RINs) impact the value of the policy's obligated parties (refiners). To accomplish this, we implement two different reduced-form methods, which allow us to avoid imposing any particular causal channel for how the policy might impact firms' value.

@@

Outside of the academic pass-through literature, some stakeholders in RFS debates express concerns that larger refiners use the RFS to disadvantage smaller ones. Refiners in the United States can be split into ``merchant" refiners, who do not blend their own fuel and are generally smaller, and ``integrated" refiners, who do. Though refineries are the parties obligated by the RFS, RINs are not actually separated from biofuels until they are blended with gasoline. All this can be seen in Figure \ref{supplychain}, with integrated refineries owning both ``Oil Refinery" and ``Fuel Blender" assets and merchant refiners owning only the former. Not owning blending assets leaves merchant refiners in the position of having to buy RINs on the market rather than being able to generate them themselves. Merchant refiners and their advocacy organizations often claim that being unable to generate RINs puts their operations at a competitive disadvantage and allows integrated refineries to sell excess RINs for windfall profits \citep[see discussion and footnotes in][p. 21-31]{EnvironmentalProtectionAgency2017}. The Environmental Protection Agency has dismissed such arguments, pointing primarily to economic research on RIN pass-through in doing so \citep{EnvironmentalProtectionAgency2017}. Research by \cite{Babcock2016} provides a theoretical explanation for why the RFS should not impact merchant and integrated refiners differentially.

We directly examine the impact of RIN price fluctuations on the stock prices of obligated refineries. First, following \cite{Lade2018a}, who used event studies to examine how RIN price shocks affect commodity markets and biorefinery stocks, we use unanticipated regulatory announcements that drastically affected the price of RINs to identify impacts on every firm in our sample, as well as selected subgroups of firms. The two large shocks are plausibly exogenous and allow us to examine a direct response in the stock price of regulated firms. Second, we fit bivariate time series models for every possible firm $\times$ RIN combination. Modeling each firm separately allows us to investigate heterogeneity among firms. While these models provide a picture of how RINs and firm stock prices are associated over time, they are subject to endogeneity problems---both RINs and refining stocks are structurally related to commodity prices, fuel demand factors, and other variables. The intent of this paper is not to identify these structural relationships, and building them into the model is beyond the scope of this research. Instead, we use these models to cross-validate and motivate our event study models, recognizing that endogeneity issues inescapably prevent us from causally interpreting the time series results, as well as to provide a graphical idea of how shocks affect firm value using impulse response functions.

We find that when RIN prices rise, the stock prices of refineries with large market capitalizations drop with a 3-5 day lag. The effect is statistically significant though economically small. Medium and small firms, however, exhibit no reaction to RIN price changes. These results are consistent across both the event study and bivariate time series analyses. Due to the reduced-form nature of our estimates, we are unable to identify any particular causal mechanism for these results, but we conclude with a list of potential hypotheses worth investigating in future research. In any case, our findings cast doubt on claims that the RFS enables integrated refiners to take advantage of merchant refiners and suggests that the magnitude of so-called RIN ``speculation" is overblown.

@@

The Renewable Fuel Standard is a nested mandate, meaning that blending higher-level biofuels also works to meet the mandate requirements at lower levels. RINs coming from corn ethanol generate D6 RINs, which can serve to fulfill only the lowest level of the mandate. Ethanol from more ``advanced" sources such as sugarcane generates advanced ethanol RINs (D5) and constitutes a smaller, nested mandate. RINs from biomass-based diesel (D4), cellulosic biofuel (D3), or cellulosic diesel (D7) fulfill the mandate in their own categories, the advanced mandate, and the total mandate simultaneously. This nested structure is visualized in Figure \ref{RINstructure}.

@@

The absence of strictness to that inequality is not merely theoretical. Federal regulation prevents most consumer fuels (excepting E85 and E15, which contain up to 85\% and 15\$ ethanol respectively, are available at a limited number of fueling stations, and can be used only in certain vehicles) from containing more than 10\% ethanol. When national gasoline stocks are saturated with 10\% ethanol, sales of additional ethanol can occur only through comparatively miniscule E85 and E15 channels. Thus, at mandate levels beyond 10\% of nationwide gasoline sales, refiners must take advantage of the nested mandate structure and sell additional biodiesel to meet their requirements \citep{Korting2019}. In those market conditions, D6 prices track closely to D4 prices \citep{Irwin2014}.

Congress intentionally set statutory RFS volume mandates optimistically high. To prevent undue financial pressures on the transportation fuels industry, Congress explicitly allows the Environmental Protection Agency (EPA) to review the statutory volumetric standards and reduce them if compliance would be infeasible. The EPA has found it necessary to invoke that power numerous times. The following statement, released along with a proposed adjustment of the 2014-2016 mandates, illuminates the EPA's role in tempering the statutory requirements:

@@

In reviewing and adjusting the yearly mandates, the EPA issues a proposed rule, gathers public comments on that proposal, and then issues a final rule. The final rule is supposed to be complete by November 30 of the preceding year (e.g., 2015's final rule should be issued by November 30, 2014). In the lifespan of the RFS, the EPA has repeatedly missed that deadline \citep{Bracmort2015}. Final rules are often made partway through the compliance year and in one case a final rule was set almost a full year after the compliance year had passed. Research by \cite{Lade2018a} demonstrates that such announcements shock RIN values as well as some commodity markets and biorefinery firm values, but no currently published research uses the shocks to identify the RFS's impact on the industry it actually regulates.

@@

While the RFS mandate's point of compliance generally rests with refiners, the associated costs can be pushed up or downstream if the industrial organization of markets allow it. As mentioned in the introduction, a developed literature discusses this very question. Most papers characterize the RFS as a subsidy to high-ethanol fuels and measure how changes in the value of the RIN 'subsidy' percolate to prices of E10 and E85 fuels. The general thrust of the literature concludes that pass-through is complete (or even more than complete) at the wholesale level \citep{Burkhardt2016,Knittel2017,Lade2019}, complete for E10 at the retail level \citep{Pouliot2017,Lade2019,Li2019}, and less than complete for most E85 \citep{Lade2019,Li2019}. The general interpretation of these results is that the RFS is irrelevant to the refiner, costly to most consumers, and beneficial to consumers of high-ethanol fuel and some E85 retailers.

Understanding pass-through does get us most of the way to understanding how the RFS affects refiners financially, but pass-through is not the only avenue by which the RFS could impose compliance costs. Costs associated with biofuel procurement and profit margins on gasoline sales lost to ethanol are just two additional potential ways the RFS could impose costs on refiners, even with complete pass-through. The methods undertaken in this paper impose no particular cost channel on examining the RFS's impact, instead electing to allow efficient financial markets price these anticipated costs. Examining stock price responses to variations in RIN prices gives a highly reduced-form but generalized and unstructured understanding of how the RFS impacts refineries.

We also analyze heterogeneities among refinery responses, rather than taking refiners as a homogeneous group. Existing research on whether the impact of the RFS is heterogeneous across obligated firms is scant. There are at least a few reasons differences might exist. For example, firms which own refining capacity in close proximity to ethanol production may react less negatively to RFS cost hikes. The RIN market is designed much like pollution permits in market-based environmental policies, so that the theoretical market equilibrium in RINs should equalize the shadow prices of additional ethanol blending across firms \citep{Montgomery1972}. Whether the RIN system operates efficiently enough to effectively accomplish that is an open empirical question. Findings by \cite{LaRiviere2017} indicate that the RFS imposed heavier burdens on retail gasoline consumers distant from ethanol production centers because ethanol's physical properties make it more expensive to transport, but this is further downstream than refining and blending.

We also investigate firm size heterogeneities by asking whether larger firms have an easier time passing-through or otherwise dealing with RFS costs. Built-in exemptions for small refiners facing economic hardship seem to indicate that the drafters of the RFS policy must have worried about differential impacts. Beyond minimizing impacts on smaller refineries, exemptions can also shift the burden of meeting the volumetric mandates to larger refineries if they are issued before the final rule \citep{Coppess2017}. Exemptions hit a low point in 2015, with only 7 petitions granted representing 3 billion gallons of gasoline, as compared to 29 petitions granted representing over 13 billion gallons just two years later \citep{EnvironmentalProtectionAgency2018}. Of course, political connections, easier access to credit, and other factors benefiting large firms may mean that smaller firms still bear the brunt of the regulation.

@@

As mentioned previously, the EPA sometimes misses regulatory announcement deadlines, resulting in major regulatory announcements that occur mid-compliance year. In 2013, the EPA released the final rule on August 6, leaked a draft proposal for 2014 on October 11, and released 2014's official proposed rule on November 15. \cite{Lade2018a} measures how 2013's mid-year announcements influenced RIN prices themselves, related commodity markets, and the stock prices of biorefining firms.  They find that those shocks, which drove the prices of all RIN varieties downward, mainly impacted commodities and firms tied to advanced biofuels, which they considered to be the marginal compliance fuel. The year 2013 was a volatile time for RIN markets, and visual inspection of RIN price histories throughout 2013 in Figure \ref{2013pricehistory} reveals that the announcement dates analyzed by \cite{Lade2018a}, while significant, are unremarkable compared to baseline volatility.

%\begin{figure}[h]
%	\caption{Price Histories for 2013 Vintage RINs}
%	\label{2013pricehistory}
%	\includegraphics[width=\textwidth]{Figures/2013RINs.pdf}
%\end{figure}

Unlike 2013, 2015's proposed and final rules, which were released on May 29 and November 30, are clearly and unambiguously apparent by visual inspection of Figure \ref{2015pricehistory}.\footnote{In an online supplement, \cite{Lade2018a} repeat the portion of their analysis detailing how policy announcements affect RIN prices for 2015 event dates. Unlike their 2013 analysis, they do not examine how 2015 announcements affected biofuel stocks or commodity markets.} Surrounding a proposed rulemaking on May 29, 2015, the price of a D6 RIN dropped from \$0.69 one week before to \$0.3775 one week afterwards, a drop of more than 45\%. A jump of similar magnitude surrounded a final rulemaking on November 30, 2015.

\begin{figure}[h]
	\caption{Price History for 2015 Vintage RINs}
	\label{2015pricehistory}
	\includegraphics[width=\textwidth]{figures/2015RINs}
\end{figure}

In this study, we focus on 2015 because the EPA made two major policy announcements that we will exploit as structural breaks following \cite{Lade2018a}. We choose 2015 over 2013, the primary year analyzed by \cite{Lade2018a}, for two reasons: First, the announcement breaks are much clearer in the data; that can be seen by comparing Figures \ref{2013pricehistory} and \ref{2015pricehistory}, but note the change in the extent of the price axis. Second, RIN prices are sufficiently high throughout 2015 to guarantee that the mandate was binding, whereas early 2013 prices were so low that the RFS may not have actually been stringent enough to alter refiner behavior \citep{Whistance2016}.

@@

This paper employs two methodologies to study the impact of the Renewable Fuel Standard on the refining industry. First, we use multivariate time series methods to quantify the response of refining firm values to shocks in RIN price series. In particular, for every Firm $\times$ RIN combination, we estimate impulse response functions using vector auto-regressions and vector error correction models. These allow us to determine the dynamic response of the value of refining firms to shocks in the value of RIN prices. This portion of our analysis builds on past literature using time series methods to unravel how RINs transmit to wholesale and consumer fuels \citep{Knittel2017} and biofuel-related commodities \citep{Whistance2014, Whistance2016}. Second, we implement an event study methodology similar to that used by \cite{Lade2018a}. Even if endogeneity complicates measurement of direct industry impact using conventional multivariate time series methods, we anticipate that the large shocks constituting nearly half of RIN prices should induce measurable impacts on refining firms if indeed there is an impact. The event study methodology also side-steps nonlinearities in the RIN-stock system that may complicate the bivariate models \citep{Serra2011}.

Both methodologies are reduced-form. By examining stock prices rather than specific product prices, we should be able to identify the impact of \textit{any} channel whereby a larger RFS mandate influences the net values of refining firms in the short-run. The downside of this approach is that we cannot validate any particular channel of causality. Nevertheless, establishing the existence or non-existence of an effect in and of itself offers insights as to whether cost pass-through is an adequate stopping point for research seeking to understand how regulated industries respond to the RFS and other fuel blend regulations.

@@

The mandate for D3 RINs is consistently minuscule, and markets for those RINs are commensurately quite thin. Therefore, we restrict our analysis to D6, D5, and D4 RINs. RIN series also differ by their year of creation; apart from some flexibility allowing limited inter-year banking and borrowing of RIN stocks, RINs are mostly used to comply with the mandate for the year in which they were generated. Thus, we limit our analysis to RINs of a single compliance year.\footnote{The majority of the literature examining RIN prices do not restrict their analyses to a single compliance year. It is not always clear how this literature stitches together price series for RINs belonging to difference compliance years, especially when RINs for multiple compliance years are sold contemporaneously.} As discussed in Section \ref{sec_2015shocks}, we analyze data from 2015. Because of some pre-year trading and the fact that RINs are not actually submitted for compliance until a few months after year end, our data does extend a little beyond a single calendar year.

All data is daily, excluding weekend and other financial market closures such as holidays. Data for RIN series come from the Oil Price Information Service (OPIS), a company that provides pricing for numerous petroleum products and is a frequent supplier of RIN data for economic researchers. Stock prices for nearly all firms are adjusted prices from Yahoo! Finance (accessed via the quantmod package in R). Stocks delisted since 2015 are not available on Yahoo! Finance and are instead sourced from Bloomberg. While these prices are not adjusted, none of those firms underwent stock splits or reverse-splits in the study period, so this should be of little consequence.

\subsection{Firm Characteristics}\label{sec_firmcharacter}
This study encompasses all publicly traded firms with at least 200,000 barrels per day of refining capacity as of January 1, 2015, plus Western Refining \citep{EnergyInformationAdministration2015}. We sort firms along two dimensions: size and exposure to RFS regulatory costs as of the beginning of 2015. We use market capitalization as a relevant measure of firm size \citep{Fama1992} and the percentage of a firm's refining capacity in PADD 2 and Alaska as our measure of exposure (or, rather, non-exposure).\footnote{PADD stands for Petroleum Administration for Defense Districts. Using refining capacity in PADDs 2, 3, and Alaska or, inversely, using refining capacity in PADDs 1 and 5 minus Alaska, hardly changes the groupings.}

PADDs are collections of states by which many government-issued petroleum data sources are aggregated. The Midwest, encompassed by PADD 2, contains the vast majority of ethanol refining capacity because refinery location decisions are driven primarily by access to feedstocks \citep{Lambert2008}. Because ethanol cannot be transported using existing pipeline infrastructure, it is plausible to think that the RFS puts refineries located near biorefineries at a comparative advantage. Geographic disparity in the impacts of the RFS have already been documented for retail gasoline prices \citep{LaRiviere2017} and pass-through of RIN prices by rack sellers at fuel terminals \citep{Pouliot2017}. Alaskan refining capacity is also included in this measure because Alaska is exempted from the RFS. We sort firms into one of nine bins, one for each element of the cross product of three market capitalization and three RFS exposure bins, as shown in Table \ref{bins}. The cutoffs were chosen somewhat arbitrarily to result in balanced bins, but, as the next paragraph further explains, the cutoffs do run parallel to important qualitative differences between firms.

@@

>>>>

\title{{\Large \bf Industry Compliance Costs Under the Renewable Fuel Standard: Evidence from Compliance Credits\thanks{This paper has benefited from comments provided by William Shughart, Ben Blau, Megan Hansen, and two anonymous referees. The Center for Growth and Opportunity at Utah State University provided financial support for data procurement and research. All errors are the author's own.}}}
\author{Arthur R. Wardle\footnote{
PhD Student, Department of Agricultural \& Resource Economics, University of California, Berkeley. \href{mailto:arw@berkeley.edu}{\tt arw@berkeley.ed} (corresponding author).}\\
Sherzod B. Akhundjanov\footnote{Assistant Professor, Department of Applied Economics, Utah State University. \href{matilto:sherzod.akhundjanov@usu.edu}{\tt sherzod.akhundjanov@usu.edu}}}

@@

The Renewable Fuel Standard (RFS), which requires oil refineries to blend ethanol into domestic fuel supplies, is a market-based policy that implements tradable compliance credits (called RINs) to better equalize compliance costs across firms. I use variation in the prices of these credits to retrieve reduced-form estimates of how the RFS impacts the stock prices of refining firms. I take advantage of unanticipated regulatory announcements that caused major swings in RIN prices to implement an event study around those announcement dates and also fit bivariate time series models to firm $\times$ RIN pairs. Both models show no significant stock price response among small or medium size firms and a small but statistically significant price response among large refiners. These findings are relevant to policy in that they cast doubt on concerns that the RFS allows integrated refiners to abuse merchant refiners. Our findings also shed light on the necessity of Small Refinery Exemptions, which are intended to shield small, financially vulnerable refiners from RFS compliance costs.
\newline

@@

The Renewable Fuel Standard (RFS), created under the Energy Policy Act of 2005 and greatly expanded by the Energy Independence and Security Act of 2007, mandates the use of various biofuels in domestic transportation fuels.\footnote{A complete description of how the RFS works is available in \cite{Schnepf2013}.} The statute itself includes volumetric mandates for cellulosic, biomass-based biodiesel, ``advanced," and total renewable fuels through 2022. Obligated parties (fuel refiners and importers) are required to submit specified numbers of Renewable Identification Numbers (RINs) to comply with the standard. RINs are created by biorefineries, who generate them whenever they produce a gallon of biofuel and link them to those gallons. The RIN is then split from those gallons upon fuel blending and becomes available for use. RINs can be traded, allowing obligated parties to comply with RFS mandates either by blending renewable fuels themselves or buying excess RINs from other parties. The construction of the RIN market is functionally equivalent to other intensity standard credit trading schemes like Renewable Portfolio Standards and similar to other market-based environmental regulations such as pollution permits.

Because regulated firms can comply with the RFS either by blending additional biofuels or by buying RINs, the basic fundamental value of a RIN (with some complications, described later in this paper) is the marginal cost to the refining sector of blending an additional unit of biofuel over a unit of gasoline.

Even high RIN prices do not necessarily impact the bottom line of obligated parties if these costs are easy to pass through to consumers and demand for transportation fuels is inelastic. Indeed, a large literature establishes that refiners are able to fully (or even more than fully) pass RFS compliance costs onto consumers \citep{Knittel2017,Pouliot2017,Burkhardt2019,Li2019,Lade2019}. This, however, does not imply that the RFS has no financial impact on US oil refiners whatsoever. Complete pass-through does not replace the profit margins on refined crude oil that counterfactually would  have been sold without the mandate and may impose infrastructure costs to accommodate ethanol blends, negotiating costs with biorefineries, and myriad costs of doing business not fully captured in models of cost pass-through. So long as fuel supply and gasoline demand are not perfectly inelastic, higher mandates must cut gasoline's market share of fuel \citep{deGorter2009}. \cite{KnittelSmith2015} give a fuller description of ethanol's impact on oil refining profitability. The goal of our analysis is to establish how changes in the prices of RFS compliance credits (RINs) impact the value of the policy's obligated parties (refiners). To accomplish this, we implement two different reduced-form methods, which allow us to avoid imposing any particular causal channel for how the policy might impact firms' value.

@@

Outside of the academic pass-through literature, some stakeholders in RFS debates express concerns that larger refiners use the RFS to disadvantage smaller ones. Refiners in the United States can be split into ``merchant" refiners, who do not blend their own fuel and are generally smaller, and ``integrated" refiners, who do. Though refineries are the parties obligated by the RFS, RINs are not actually separated from biofuels until they are blended with gasoline. All this can be seen in Figure \ref{supplychain}, with integrated refineries owning both ``Oil Refinery" and ``Fuel Blender" assets and merchant refiners owning only the former.\footnote{Note that refining firms can also own biorefineries, as Valero does.}  Not owning blending assets leaves merchant refiners in the position of having to buy RINs on the market rather than being able to generate them themselves. Merchant refiners and their advocacy organizations often claim that being unable to generate RINs puts their operations at a competitive disadvantage and allows integrated refineries to sell excess RINs for windfall profits \citep[see discussion and footnotes in][p. 21-31]{EnvironmentalProtectionAgency2017}. The Environmental Protection Agency has dismissed such arguments, pointing primarily to economic research on RIN pass-through in doing so \citep{EnvironmentalProtectionAgency2017}. Research by \cite{Babcock2016} provides a theoretical explanation for why the RFS should not impact merchant and integrated refiners differentially.

Through the Small Refinery Exemption (SRE) system, refiners undergoing economic hardship are able to apply for waivers to the Renewable Fuel Standard. Usage of SREs was at a low point in 2015, but the Trump administration's widespread usage of SREs has become a flashpoint in RFS politics. How and whether the RFS affects firms differentially is important to understanding the legitimacy of the SRE system.

We directly examine the impact of RIN price fluctuations on the stock prices of obligated refineries. First, following \cite{Lade2018a}, who used event studies to examine how RIN price shocks affect commodity markets and biorefinery stocks, we use unanticipated regulatory announcements that drastically affected the price of RINs to identify impacts on every firm in our sample, as well as selected subgroups of firms. The two large shocks are plausibly exogenous and large in magnitude, allowing us to examine a direct response in the stock price of regulated firms. Second, we fit bivariate time series models for every possible firm $\times$ RIN combination. Modeling each firm separately allows us to investigate heterogeneity among firms. While these models provide a picture of how RINs and firm stock prices are associated over time, they are subject to endogeneity problems---both RINs and refining stocks are structurally related to commodity prices, fuel demand factors, and other variables. The intent of this paper is not to identify these structural relationships, and building them into the model is beyond the scope of this research. Instead, we use these models to validate and motivate our event study models, recognizing that endogeneity issues inescapably prevent us from causally interpreting the time series results, as well as to provide a graphical idea of how shocks affect firm value using impulse response functions.

We find that when RIN prices rise, the stock prices of refineries with large market capitalizations drop with a 3-5 day lag. The effect is statistically significant though economically small. Medium and small firms, however, exhibit no reaction to RIN price changes. These results are consistent across both the event study and bivariate time series analyses. Due to the reduced-form nature of our estimates, we are unable to identify any particular causal mechanism for these results, but we conclude with a list of potential hypotheses worth investigating in future research. In any case, our findings cast doubt on claims that the RFS enables integrated refiners to take advantage of merchant refiners and questions the legitimacy of increasingly widespread Small Refinery Exemption issuance.

@@

The Renewable Fuel Standard is a nested mandate, meaning that blending higher-level biofuels also works to meet the mandate requirements at lower levels (i.e. the entire mandate could theoretically be met using only biomass-based diesel and cellulosic biofuel, but this would be excessively expensive). RINs coming from corn ethanol generate D6 RINs, which can serve to fulfill only the lowest level of the mandate. Blending corn ethanol is generally cheaper than blending any other biofuel, so refiners tend to use D6 RINs to comply with as much of the mandate as possible. Ethanol from more ``advanced" sources such as sugarcane generates advanced ethanol RINs (D5) and constitutes a smaller, nested mandate. RINs from biomass-based diesel (D4), cellulosic biofuel (D3), or cellulosic diesel (D7) fulfill the mandate in their own categories, the advanced mandate, and the total mandate simultaneously. This nested structure is visualized in Figure \ref{RINstructure}.

@@

The absence of strictness to that inequality is not merely theoretical. Federal regulation prevents most consumer fuels (excepting E85 and E15, which contain up to 85\% and 15\% ethanol respectively, are available at a limited number of fueling stations, and can be used only in certain vehicles) from containing more than 10\% ethanol. When national gasoline stocks are saturated with 10\% ethanol, sales of additional ethanol can occur only through comparatively miniscule E85 and E15 channels. Thus, at mandate levels beyond 10\% of nationwide gasoline sales, refiners must take advantage of the nested mandate structure and sell additional biodiesel to meet their requirements \citep{Korting2019}. In those market conditions, D6 prices track closely to D4 prices \citep{Irwin2014}.

Congress intentionally set statutory RFS volume mandates optimistically high. To prevent undue financial pressures on the transportation fuels industry, Congress explicitly allows the Environmental Protection Agency (EPA) to review the statutory volumetric standards and reduce them if compliance would be infeasible for suppliers. The EPA has found it necessary to invoke that power numerous times. The following statement, released along with a proposed adjustment of the 2014-2016 mandates, illuminates the EPA's role in tempering the statutory requirements:

@@

In reviewing and adjusting the yearly mandates, the EPA issues a proposed rule, gathers public comments on that proposal, and then issues a final rule. The final rule is supposed to be complete by November 30 of the preceding year (e.g., 2015's final rule should be issued by November 30, 2014). In the lifespan of the RFS, the EPA has repeatedly missed that deadline \citep{Bracmort2015}. Final rules are often made partway through the compliance year and in one case a final rule was set almost a full year after the compliance year had passed. Research by \cite{Lade2018a} demonstrates that such announcements shock RIN values as well as some commodity markets and biorefinery firm values, but no currently published research uses the shocks to identify the RFS's impact on the value of the firms it actually regulates.

@@

While the RFS mandate's point of compliance generally rests with refiners, its incidence can be pushed up or downstream if the industrial organization of markets allow it. As mentioned in the introduction, a developed literature discusses this very question. Most papers characterize the RFS as a subsidy to high-ethanol fuels and measure how changes in the value of the RIN `subsidy' percolate to prices of E10 and E85 fuels. The general thrust of the literature concludes that pass-through is complete (or even more than complete) at the wholesale level \citep{Knittel2017,Burkhardt2019,Lade2019}, complete for E10 at the retail level \citep{Pouliot2017,Lade2019,Li2019}, and less than complete for most E85 \citep{Lade2019,Li2019}. The general interpretation of these results is that the RFS is on average irrelevant to the refiner, costly to most consumers, and beneficial to consumers of high-ethanol fuel and some E85 retailers.

Understanding pass-through does get us most of the way to understanding how the RFS affects refiners financially, but pass-through is not the only avenue by which the RFS could impose compliance costs. Profit margins on gasoline sales lost to ethanol and costs associated with biofuel procurement are just two additional potential ways the RFS could impose costs on refiners, even at the margin of the current mandate and with complete pass-through. If positive shocks to the RIN market are unanticipated by investors, stock prices should adjust post-shock to reflect investor's revised expectations for future profits. The methods undertaken in this paper impose no particular cost channel on examining the RFS's impact, instead electing to allow financial markets price these anticipated costs. Examining stock price responses to variations in RIN prices gives a highly reduced-form but generalized and unstructured understanding of how the RFS impacts refineries.

We also analyze whether a refiner's size affects their response to mandate changes, rather than taking refiners as a homogeneous group. Existing research on whether the impact of the RFS is heterogeneous across obligated firms is scant. The RIN market is designed much like pollution permits in market-based environmental policies, so that the theoretical market equilibrium in RINs should equalize the shadow prices of additional ethanol blending across firms \citep{Montgomery1972}. Built-in exemptions for small refiners facing economic hardship (SREs) seem to indicate that the drafters of the RFS policy must have worried about differential impacts. Beyond minimizing impacts on smaller refineries, exemptions can also shift the burden of meeting the volumetric mandates to larger refineries if they are issued before the final rule \citep{Coppess2017}. Exemptions hit a low point in 2015, with only 7 petitions granted representing 3 billion gallons of gasoline, as compared to 29 petitions granted representing over 13 billion gallons just two years later \citep{EnvironmentalProtectionAgency2018}. Of course, political connections, easier access to credit, and other factors benefiting large firms may mean that smaller firms still bear the brunt of the regulation.

@@

As mentioned previously, the EPA sometimes misses regulatory announcement deadlines, resulting in major regulatory announcements that occur mid-compliance year. In 2013, the EPA released the final rule on August 6, leaked a draft proposal for 2014 on October 11, and released 2014's official proposed rule on November 15. \cite{Lade2018a} measures how 2013's mid-year announcements influenced RIN prices themselves, related commodity markets, and the stock prices of biorefining firms.  They find that those shocks, which drove the prices of all RIN varieties downward, mainly impacted commodities and firms tied to advanced biofuels, which they considered to be the marginal compliance fuel. The year 2013 was a volatile time for RIN markets, and visual inspection of RIN price histories throughout 2013 in Figure \ref{2013pricehistory} reveals that the announcement dates analyzed by \cite{Lade2018a}, while significant, are unremarkable compared to volatility throughout the year.

\begin{figure}[h]
	\caption{Price Histories for 2013 Vintage RINs}
	\label{2013pricehistory}
	\includegraphics{figures/2013RINs.pdf}
\end{figure}

Unlike 2013, 2015's proposed and final rules, which were released on May 29 and November 30, are clearly and unambiguously apparent by visual inspection of Figure \ref{2015pricehistory}.\footnote{In an online supplement, \cite{Lade2018a} repeat the portion of their analysis detailing how policy announcements affect RIN prices for 2015 event dates. Unlike their 2013 analysis, they do not examine how 2015 announcements affected biofuel stocks or commodity markets.} The mandate modifications in the May 29, 2015 proposed rule were lower than expected; the price of D6 RINs dropped from \$0.69 one week before to \$0.3775 one week afterwards, a drop of more than 45\%. The final rule pushed the mandate back up, and a jump of similar magnitude to the original plummet surrounded that rule's announcement on November 30, 2015.

\begin{figure}[h]
	\caption{Price History for 2015 Vintage RINs}
	\label{2015pricehistory}
	\includegraphics{figures/2015RINs}
\end{figure}

In this study, we focus on 2015 because the EPA made two major policy announcements that we will exploit as structural breaks following \cite{Lade2018a}. We choose 2015 over 2013, the primary year analyzed by \cite{Lade2018a}, for two reasons: First, the announcement breaks are much clearer in the data; that can be seen by comparing Figures \ref{2013pricehistory} and \ref{2015pricehistory}, noting the change in the extent of the price axis. Second, RIN prices are sufficiently high throughout 2015 to guarantee that the mandate was binding, whereas early 2013 prices were so low that the RFS may not have actually been stringent enough to alter refiner behavior \citep{Whistance2016}.

@@

This paper employs two methodologies to study the impact of the Renewable Fuel Standard on the refining industry. First, we implement an event study methodology similar to that used by \cite{Lade2018a}. The steep and sudden swings present in the 2015 RIN price series suggest that investors did not foresee the magnitude of EPA changes to the RFS mandate ahead of time. We anticipate that the large shocks constituting nearly half the original RIN prices should induce measurable impacts on refining firms if indeed there is an impact. The event study methodology mitigates the issues facing traditional time series analysis like endogeneity and the complexity of structural relationships in commodity price systems. For example, this method obviates the need to explicitly account for nonlinearities in the RIN-stock system that may complicate a typical time series model \citep{Serra2011}.

For completeness, we also use multivariate time series methods to quantify the response of refining firm values to variations in RIN price series. In particular, for every Firm $\times$ RIN combination, we estimate impulse response functions using vector auto-regression models. These allow us to determine the dynamic response of the value of refining firms to shocks in the value of RIN prices. This portion of our analysis builds on past literature using time series methods to unravel how RINs transmit to wholesale and consumer fuels \citep{Knittel2017} and biofuel-related commodities \citep{Whistance2014, Whistance2016}. We perceive these models as being a secondary gut-check to the headline event study results.

Both methodologies are reduced-form. By examining stock prices rather than specific product prices, we should be able to identify the impact of \textit{any} channel whereby changes in the expected cost of RFS compliance influence the net values of refining firms in the short-run. The downside of this approach is that we cannot validate any particular channel of causality. Nevertheless, establishing the existence or non-existence of an effect in and of itself offers insights as to whether cost pass-through is an adequate stopping point for research seeking to understand how regulated industries respond to the RFS and other fuel blend regulations.

@@

The mandate for D3 RINs is consistently minuscule, and markets for those RINs are commensurately quite thin. Therefore, we restrict our analysis to D6, D5, and D4 RINs. RIN series also differ by their year of creation; apart from some flexibility allowing limited inter-year banking and borrowing of RIN stocks, RINs are mostly used to comply with the mandate for the year in which they were generated. Thus, we limit our analysis to RINs of a single compliance year.\footnote{The majority of the literature examining RIN prices do not restrict their analyses to a single compliance year. It is not always clear how this literature stitches together price series for RINs belonging to difference compliance years, especially when RINs for multiple compliance years are sold contemporaneously. One simple approach, taking the average of all traded prices for a given RIN variety, requires imposing the additional assumption that banking/borrowing constraints are non-binding.} As discussed in Section \ref{sec_2015shocks}, we analyze data from 2015. Because of some pre-year trading and the fact that RINs are not actually submitted for compliance until a few months after year end, our data does extend a little beyond a single calendar year.

All data is daily, excluding weekend and other financial market closures such as holidays. Data for RIN series come from the Oil Price Information Service (OPIS), a company that provides pricing for numerous petroleum products and is a frequent supplier of RIN data for economic researchers. Adjusted stock prices as well as the Russel 3000 index and WTI oil price index come from Finaeon Global Financial Data.

\subsection{Firm Characteristics}\label{sec_firmcharacter}
This study encompasses all publicly traded firms with at least 200,000 barrels per day of refining capacity as of January 1, 2015, plus Western Refining \citep{EnergyInformationAdministration2015}. Below this cutoff, fewer firms are publicly traded and the publicly traded ones tend to be companies whose main industry is not fuels (e.g. Delta Airlines). We use market capitalization as a relevant measure of firm size \citep{Fama1992}, sorting firms into one of three bins as seen in figure \ref{bins}. The cutoffs were chosen somewhat arbitrarily to result in balanced bins, but, as the next paragraph further explains, the cutoffs do run parallel to important qualitative differences between firms.

@@

\section{Appendix}

\subsection{Bivariate Time Series Models Forced to Six Lags}
